2.1 APL Data Quality Check with additonal cleaning

### 2024-02-13 2_scripts/2.1 APL Data Quality Check.qmd
Updated APL data cleaning code, match for quintmat & quintsoc info

table(is.na(RFD4682_e$`N-IgG_RESULT`)==F,
      is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F, useNA = 'ifany')
```
Some additional cleaning/editing, making anti-S quantitative
```{r}
# Missing values
summary(is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F) 
# N=36, beware of "not available" values, better recode those as NAs

@@

# index
index<-RFD4682_e$`RBD-IgGII_RESULT` %>%
  grepl("index", .,fixed = F, ignore.case = T)
table(index) # N=104946
```
#### Yuan found this: anti-N is only available for the Omicron stage
```{r}
# The APL Data:
RFD4682_e$COLLECTION_DATE %>%
  summary()

@@

# Anti-S assay results is available to most of the participants
RFD4682_e[is.na(RFD4682_e$clean_IDe)==F &
          is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F,]$COLLECTION_DATE %>%
  summary()
```
#### A quick clean of the Anti-N data
```{r}
# Remove unecessary 'index'
table(RFD4682_e$`N-IgG_RESULT`)

@@

#### Add a year-month var
```{r}
# only keep the month, remove the data
RFD4682_e$Col_Month<-format(as.Date(RFD4682_e$COLLECTION_DATE), "%Y-%m")
table(RFD4682_e$Col_Month, useNA='ifany')

@@

#### Save a cleaned ver of RFD4682e
```{r}
save(RFD4682_e, file = "../1_data/private/RFD4682e.RData")
```

>>>>

table(is.na(RFD4682_e$`N-IgG_RESULT`)==F,
      is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F, useNA = 'ifany')
```

Some additional cleaning/editing, making anti-S quantitative

```{r}
# Missing values
summary(is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F) 
# N=36, beware of "not available" values, better recode those as NAs

@@

# index
index<-RFD4682_e$`RBD-IgGII_RESULT` %>%
  grepl("index", .,fixed = F, ignore.case = T)
table(index) # N=104946
```

#### Yuan found this: anti-N is only available for the Omicron stage

```{r}
# The APL Data:
RFD4682_e$COLLECTION_DATE %>%
  summary()

@@

# Anti-S assay results is available to most of the participants
RFD4682_e[is.na(RFD4682_e$clean_IDe)==F &
          is.na(RFD4682_e$`RBD-IgGII_RESULT`)==F,]$COLLECTION_DATE %>%
  summary()
```

#### As required by Matthew, add quintmat & qunitsoc variable
```{r}
# load a matching scale, by 6-digit postal codes
pcode_match <- read.delim("~/COVID-assay-adjust-methods/1_data/private/ot_ccp_postalcodecodeequivalence.txt")

# load updated version of APL
library(readxl)
RFD4682_v2 <- read_excel("1_data/private/RFD4682_Data_for release Jan 2024.xlsx", 
    sheet = "RFD4682_Full_Data", col_types = c("text", 
        "numeric", "text", "numeric", "numeric", 
        "date", "text", "text", "date", "date", 
        "text", "numeric", "numeric", "text", 
        "text", "text", "text", "text", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))
View(RFD4682_v2)

# match 6-digit postal codes to the first ver
RFD4682_e$PAT_POSTALCODE <- RFD4682_v2$PAT_POSTALCODE

# cleanup NA
length(RFD4682_e$PAT_POSTALCODE[RFD4682_e$PAT_POSTALCODE=="N/A"]) # needs to cleanup "N/A"
RFD4682_e$PAT_POSTALCODE[RFD4682_e$PAT_POSTALCODE=="N/A"]<-NA
table(is.na(RFD4682_e$PAT_POSTALCODE))

# does the postal code match?
# create a truncated ver of 6-digit postal code
RFD4682_e$PAT3<-substr(RFD4682_e$PAT_POSTALCODE, 1, 3)

RFD4682_e$PAT_match<-if_else(RFD4682_e$PAT_FSA==RFD4682_e$PAT3, 1, 0)
table(RFD4682_e$PAT_match, useNA = "ifany") # 1065 unmatch, 3276 missing

RFD4682_e %>%
  .[.$PAT_match == 0 &  !is.na(.$PAT3),] %>%
  select(PAT_POSTALCODE, PAT3, PAT_FSA)
# N=1756 unmatched cases btw APL ver 1 & 2. Talked to M. Knight and he would like to use the updated ver

RFD4682_e<-merge(RFD4682_e, select(pcode_match, PCODE, QUINTMAT, QUINTSOC),
      by.x = "PAT_POSTALCODE", by.y="PCODE", all.x = T, all.y = F)

# a clean version of if for M.Knight
RFD4682_e1<-select(RFD4682_e, -c("PAT3", "PAT_match", "N", "Col_Month"))

#### A quick clean of the Anti-N data

```{r}
# Remove unecessary 'index'
table(RFD4682_e$`N-IgG_RESULT`)

@@

#### Add a year-month var

```{r}
# only keep the month, remove the data
RFD4682_e$Col_Month<-format(as.Date(RFD4682_e$COLLECTION_DATE), "%Y-%m")
table(RFD4682_e$Col_Month, useNA='ifany')

@@

#### Save a cleaned ver of RFD4682e

```{r}
save(RFD4682_e, file = "../1_data/private/RFD4682e.RData")

save(RFD4682_e1, file = "../1_data/private/RFD4682e1.RData")
```
